Evaluating Informative Hypotheses with Equality and Inequality Constraints:\\ A Tutorial Using the Bayes Factor via the Encompassing Prior Approach

### 2021-11-27 Documents/Paper/paper.Rnw
corrige eng

\title{Evaluating Informative Hypotheses with Equality and Inequality Constraints:\\ A tutorial using the Bayes Factor via the Encompassing Prior Approach}
\shorttitle{The Bayes Factor via the Encompassing Prior Approach}

@@

\abstract{When conducting a study, researchers usually have expectations based on hypotheses or theoretical perspectives they want to evaluate. Equality and inequality constraints on the model parameters are used to formalize researchers' expectations or theoretical perspectives into the so-called “\textit{informative hypotheses}”. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are unsuitable for testing complex informative hypotheses. An alternative approach is to use the Bayes Factor. In particular, the Bayes Factor based on the “\textit{encompassing prior approach}” allows researchers to easily evaluate complex informative hypotheses. This paper provides a detailed introduction to the Bayes Factor with encompassing prior. First, all steps and elements involved in the formalization of  informative hypotheses and computation of the Bayes Factor are described. Next, we apply this method to a real case scenario considering the attachment theory. Precisely, we analyzed the relative role of maternal and paternal attachment on children's social-emotional development by comparing the different theoretical perspectives debated in the literature.

@@

\noindent The evaluation of hypotheses is one of the main goals of empirical research. In fact, when conducting a study, researchers usually want to evaluate the plausibility of specific hypotheses according to the collected data. These hypotheses can be formalized in accordance with researchers' expectations or theoretical perspectives into the so-called “\textit{informative hypotheses}”. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are not suitable for evaluating the plausibility of complex informative hypotheses. The Bayes Factor via the “\textit{encompassing prior approach}”, instead, allows researchers to easily evaluate the plausibility of complex informative hypotheses. This paper provides a detailed introduction to the Bayes Factor with encompassing prior. First, all steps and elements involved in the formalization of  informative hypotheses and computation of the Bayes Factor are described. Next, we apply this method to a real case scenario considering the attachment theory. Precisely, we analyzed the relative role of maternal and paternal attachment on children's social-emotional development by comparing the different theoretical perspectives debated in the literature. This approach is of particular interest in applied research as it allows researchers to formalize hypotheses with great flexibility and to directly evaluate which hypothesis is the most supported by the data.
}

@@

The dominant statistical approach to evaluate research hypotheses in psychology is the Null Hypothesis Significance Testing (NHST). In the literature, however, the utility and validity of NHST are largely debated \parencite{wassersteinMovingWorld052019}. This approach presents indeed several limitations. First, the NHST places a narrow focus on statistical testing rather than on the formalization of hypotheses. This has usually led researchers to evaluate data against the catch-all null-hypothesis that nothing is going on rather than testing their specific expectation [and the alternative hypothesis is rarely formalized]. Second, the p-value does not quantify the evidence in favour of one hypothesis, thus, it is not possible to \textit{“accept”} a hypothesis but only to \textit{“reject”} it. This has some drawbacks since, in case of not significant results, the researchers are left in a state of indecision. Third, NHST does not allow testing multiple hypotheses at the same time. Using the NHST the null hypothesis is tested only against a single alternative. Fourth, NHST is unsuitable for testing broad classes of hypotheses with equality and inequality constraints \parencite{mulderSimpleBayesianTesting2019}. Expectations can be evaluated using one-side tests; however, when more groups or more variables are involved, it is not possible to evaluate complex parameter constraints that reflect researchers' expectations \parencite{vandeschootIntroductionBayesianModel2011}.

Model comparison is a different approach that allows researchers to compare multiple hypotheses and identify which is the most supported by the data \parencite{mcelreathStatisticalRethinkingBayesian2020}. Hypotheses are first formalized as statistical models according to researchers expectations or theoretical perspectives. Subsequently, it is possible to evaluate which is the most supported model among those considered according to the data. To do that a popular approach is to use information criteria such as the AIC or BIC criteria that estimate models ability to predict new data \parencite{wagenmakersAICModelSelection2004,akaike1973a, schwarzEstimatingDimensionModel1978}. Model comparison has several advantages and, in particular, it allows to directly compute the relative plausibility of each model given the data and the set of models considered. However, model comparison using information criteria is not appropriate to evaluate hypotheses that include complex parameter constraints reflecting researchers' expectations. In fact, information criteria evaluate models complexity according to the number of parameters, but they do not take into account possible order constraints on the parameters.

An alternative criterium to evaluate research hypotheses in a model comparison is the Bayes Factor. During the last 25 years, there has been an increasing interest in the Bayes Factor and its use has been proposed as the solution to the critical issues of the NHST \parencite{heckReviewApplicationsBayes2020, mulderEditorsIntroductionSpecial2016}. Although it has its own limitations and it does not solve the fundamental issues of the misuse of statistical techniques \parencite{schadWorkflowTechniquesRobust2021, gelmanBayesianDataAnalysis2013}, the Bayes Factor offers some clear advantages. In particular, the Bayes Factor allows us to compare hypotheses obtaining a relative index of evidence, like the information criteria, but it also allows us to easily compare complex research hypotheses. In fact, so-called \textit{“informative hypotheses”} (i.e., hypotheses containing information about the ordering of the model parameters) can be formalized according to researchers' expectations or theoretical perspectives using equality and inequality constraints \parencite{vandeschootIntroductionBayesianModel2011}. Subsequently, these hypotheses can be compared against each other using the Bayes Factor.

The evaluation of informative hypotheses is of particular interest to the researchers as it allows them to assess specific complex expectations and theoretical perspectives directly. In the literature, a particular approach allowing to easily compute the Bayes Factor with complex informative hypotheses has received increasing attention: \textit{“the Bayes Factor with encompassing prior”}. \textcite{vandeschootIntroductionBayesianModel2011} presented a general introduction to informative hypotheses testing using the Bayes Factor with the encompassing prior approach, whereas \textcite{hoijtinkInformativeHypothesesTheory2012} offered a more detailed description of the development of this method. Other studies, instead, considered the application of this approach with specific statistical models; for example: mixed effect models \parencite{katoBayesianApproachInequality2006}, evaluation of correlation coefficients \parencite{mulderBayesFactorsTesting2016} and multinomial models \parencite{heckMultinomialModelsLinear2019}. In addition, \textcite{guBayesianEvaluationInequality2014} proposed a general approximate procedure to evaluate inequality constraints in a wide range of statistical models.

However, all studies mentioned above considered only informative hypotheses with inequality constraints. Evaluating in the same hypothesis equality and inequality constraints, instead, it was not possible and equality constraints had to be approximated to “\textit{about equality constraints}” (i.e., a equality constraints of type $\theta_i = \theta_j$ is approximated to $|\theta_i - \theta_j| < \xi$ for a value of $\xi$ small enough).

Only recently, \textcite{guApproximatedAdjustedFractional2018} introduced an approximate procedure to evaluate both equality and inequality constraints using the Bayes Factor with the encompassing prior approach in a wide range of statistical models (i.e., generalized linear mixed models and structural equation models). Subsequently, this approach was extended by \textcite{mulderBayesFactorTesting2019} to generalized multivariate probit models and, finally, \textcite{mulderSimpleBayesianTesting2019} proposed an accurate procedure (i.e., not based on an approximation) that allows testing informative hypotheses with equality and inequality constraints in linear regression models.

The development and implementation of this approach are of particular interest because, although with its limits, it allows researchers to directly evaluate their expectations and hypotheses. The available literature, however, is rather technical. The complexity of these articles makes it difficult for researchers not familiar with this approach to clearly understand all the steps involved. On the other hand, articles offering a general introduction to the Bayes Factor do not provide enough details to allow readers to autonomously apply this method to their problems, but they usually rely on ad-hoc solutions implemented in some statistical software \parencite[for example consider][]{hoijtinkTutorialTestingHypotheses2019a}.

The aim of this paper, therefore, is to offer a clear and detailed description of the Bayes Factor with the encompassing prior approach to allow other researchers to apply this approach in their studies. In particular, we refer to the approximated method proposed by \textcite{guApproximatedAdjustedFractional2018} as it applies to a wider range of conditions than the more accurate approach proposed by \textcite{mulderSimpleBayesianTesting2019}. The paper is organized as follows. First, the method is introduced providing a detailed description of all steps and elements involved in the formalization of informative hypotheses and Bayes Factor computation. Subsequently, an application of the method to real data is presented to offer the opportunity to discuss the typical issues encountered in real complex scenarios.

@@

The evaluation of informative hypotheses with equality and/or inequality constraints involves several steps and elements. In this section, first, we describe how to formulate informative hypotheses. Subsequently, we introduce the Bayes Factor considering the encompassing prior approach based on the approximated method proposed by \textcite{guApproximatedAdjustedFractional2018}.

@@

\subsection{Formulation of Informative Hypothesis}\label{sec:informative-hypothesis}

Informative hypotheses can be defined according to researchers' expectation, evidence from the literature or theoretical perspectives and they are formed by equality and/or inequality constraints on certain model parameters. These constraints are obtained as a linear combination of certain parameters and eventual constant values. For example, it is possible to state that two parameters are equal ($\theta_i = \theta_j$), one parameter is greater than another ($\theta_i > \theta_j$), or express other complex conditions as the difference between two parameter is less than a given value ($\theta_i - \theta_j < .5$; $2\times\theta_i - \theta_j < .1 \times \sigma$).

Thus, an informative hypothesis $H_i$ with equality and inequality constraints can be expressed in the form
\begin{equation}
H_i:\ \ R_E\bm{\theta} = r_E \ ~ \ \& \ ~ \ R_I\bm{\theta} > r_I,
\end{equation}
where $R_E$ is a matrix expressing the equality constraints and $r_E$ is a vector containing the constant values of the equality constraints. Whereas, $R_I$ is a matrix expressing the inequality constraints and $r_E$ is a vector containing the constant values of the inequality constraints. Finally, $\bm{\theta}$ is a vector with the model parameters involved in the constraints.

As an example, consider a study evaluating the efficacy of a new psychological treatment that is supposed to improve a given cognitive ability. In the study, a control group receiving no treatment and another group receiving the traditional treatment were included as a comparison. Moreover, imagine that the new psychological treatment was administered in three different modalities to different groups: individually, in-group, and online. Researchers expect no differences between the three modalities but they hypothesize that the new treatment will perform better than the traditional one and this, in turn, will be better than the no-treatment control condition. We can express this hypothesis as
\begin{equation}
H_i:\ \ \theta_{control}< \theta_{traditional} < \theta_{individual} = \theta_{group} = \theta_{online},
\end{equation}
were $\theta_{control}$ is the parameter estimating the average score of the no-treatment control group, $\theta_{traditional}$ is the parameter estimating the average score of the group receiving the traditional treatment, and $\theta_{individual}$, $\theta_{group}$, and $\theta_{online}$ are the parameters estimating the average scores of the groups receiving the new treatment individually, in group, and on-line, respectively. The corresponding formulation of the hypothesis using the matrix notation introduced before is
\begin{equation}
\begin{aligned}
H_i: &\\
& R_E\bm{\theta} =
\begin{bmatrix}
0 & 0 & 1 & -1 & 0\\
0 & 0 & 0 & 1 & -1
\end{bmatrix}        \begin{bmatrix} \theta_{control} \\ \theta_{traditional} \\ \theta_{individual} \\ \theta_{group} \\ \theta_{online}\end{bmatrix} = \begin{bmatrix} 0 \\ 0\end{bmatrix} = r_E,\\~\\
& R_I\bm{\theta} =
\begin{bmatrix}
-1 & 1 & 0 & 0 & 0\\
0 & -1 & 1 & 0 & 0
\end{bmatrix}        \begin{bmatrix} \theta_{control} \\ \theta_{traditional} \\ \theta_{individual} \\ \theta_{group} \\ \theta_{online}\end{bmatrix} > \begin{bmatrix} 0 \\ 0\end{bmatrix} = r_I.
\end{aligned}
\end{equation}
Note how each row of $R_E$ and $R_I$ matrices expresses an equality or an inequality constraint, respectively. For example, in the first row of $R_E$ we have $\theta_{individual} - \theta_{group} = 0$ (i.e., $\theta_{individual} = \theta_{group}$) and in the first row of $R_I$ we have $\theta_{traditional} - \theta_{control} > 0$ (i.e., $\theta_{traditional} > \theta_{control}$).

Now that we have understood how to define an informative hypothesis with equality and inequality constraints introducing an appropriate notation, let's see how to evaluate an informative hypothesis using the Bayes Factor with the encompassing prior approach.

@@

The Bayes Factor of hypothesis $H_1$ against a competing hypothesis $H_2$ is defined as the ratio between the marginal likelihoods of the two hypotheses:
\begin{equation}
BF_{12} = \frac{Pr(Y|H_1)}{Pr(Y|H_2)} = \frac{\int l(Y|\theta_1, H_1) \pi(\theta_1| H_1)\,d\theta_1}{\int l(Y|\theta_2, H_2) \pi(\theta_2| H_2)\,d\theta_2},
\end{equation}
where $Y$ indicates the data, $\theta_i$ is the vector of parameters under the hypothesis $H_i$, $l(Y|\theta_i, , H_i)$ is the likelihood function under the hypothesis $H_i$, and $\pi(\theta_i| H_i)$ is the prior of the parameters under the hypothesis $H_i$. The marginal likelihood $Pr(Y|H_i)$ can be interpreted as a measure of the plausibility of the data under $H_i$.

Therefore, the Bayes Factor $BF_{12}$ quantifies the relative support of the data for the two competing hypotheses (and not the ratio between the probability of the two hypotheses). For values close to 1, the Bayes Factor indicates that $H_1$ and $H_2$ have similar support from the data. Larger values indicate evidence in favour of $H_1$, whereas values close to 0 indicate evidence in favour of $H_2$. The ratio between the posterior probabilities of the two hypotheses can be computed as
\begin{equation}
\frac{Pr(H_1|Y)}{Pr(H_2|Y)} = \overbrace{\frac{Pr(Y|H_1)}{Pr(Y|H_2)}}^{BF_{12}} \times \frac{Pr(H_1)}{Pr(H_2)},
\end{equation}
where $Pr(H_i)$ is the prior probability of $H_i$ and $Pr(H_i|Y)$ is the posterior probability of $H_i$. For a detailed description of the Bayes Factor considering also its interpretation and application in different contexts see \textcite{wagenmakersBayesianHypothesisTesting2010, mulderEditorsIntroductionSpecial2016, heckReviewApplicationsBayes2020}.

Note that to compute the marginal likelihood $Pr(Y|H_i)$ of a hypothesis $H_i$ it is necessary to integrate the product between the likelihood and the prior. These integrals, however, are usually difficult to compute [in particular when hypotheses include order constraints]. In the case of hypotheses with equality and inequality constraints, however, it is possible to simplify the computation of the Bayes Factor using the encompassing prior approach.

@@

The basic idea of the encompassing prior approach is to consider an informative hypothesis as a subset of the parameter space of an unconstrained model. Thus, to evaluate the plausibility of a hypothesis we can consider the proportion of parameter space of the unconstrained model that satisfies the constraints. More specifically, given an informative hypothesis $H_i$ and an unconstrained model $H_u$ (or \textit{encompassing model}) that does not contain any constraints on the parameters, if the prior under $H_i$ is defined as a truncation of the proper prior under $H_u$ (\textit{encompassing prior}) according to the constraints, then the Bayes Factor between $H_i$ and $H_u$ can be written as
\begin{equation}\label{eq:BF-encompassing}
\begin{aligned}
BF_{iu} &= \frac{Pr(\text{Inequality Const|Equality Const, Data, } H_u)}{\pi(\text{Inequality Const|Equality Const, } H_u)} \times \frac{Pr(\text{Equality Const|Data, } H_u)}{\pi(\text{Equality Const}|H_u)} \\ ~\\
&= \frac{Pr(R_I\theta > r_I | R_E\theta =r_E, Y, H_u)}{\pi(R_I\theta > r_I | R_E\theta =r_E, H_u)} \times \frac{Pr(R_E\theta =r_E| Y, H_u)}{\pi(R_E\theta =r_E| H_u)}.
\end{aligned}
\end{equation}

The first term is the ratio between the conditional posterior probability and the conditional prior probability that the inequality constraints hold under the unconstrained model $H_u$ given the equality constraints. The second term is the ratio between marginal posterior density and the marginal prior density of the equality constraints under $H_u$ \parencite[the well-known Savage–Dickey density ratio; ][]{dickeyWeightedLikelihoodRatio1971, wetzelsEncompassingPriorGeneralization2010}. In particular, the four elements can be interpreted as:
\begin{itemize}
  \item{The \textbf{conditional posterior probability} $Pr(R_I\theta > r_I | R_E\theta =r_E, Y, H_u)$ is a measure of the fit of the inequality constraints of $H_i$ under $H_u$.}
  \item{The \textbf{conditional prior probability} $\pi(R_I\theta > r_I | R_E\theta =r_E, H_u)$ is a measure of the complexity of the inequality constraints of $H_i$ under $H_u$.}
  \item{The \textbf{marginal posterior density} $Pr(R_E\theta =r_E| Y, H_u)$ is a measure of the fit of the equality constraints of $H_i$ under $H_u$.}
  \item{The \textbf{marginal prior density} $\pi(R_E\theta =r_E| H_u)$ is a measure of the complexity of the equality constraints of $H_i$ under $H_u$.}
\end{itemize}
Summarizing at the denominators we have measures of the \textbf{complexity} of the inequality and equality constraints of the informative hypothesis $H_i$; at the numerators, instead, we have measures of the \textbf{fit} of the data to the inequality and equality constraints of the informative hypothesis $H_i$. Thus, if the hypothesis $H_i$, although applying constraints to the parametric space (less complexity), is still able to provide a good description of the data, the Bayes Factor will favour $H_i$. On the contrary, if the support of the data is poor, the Bayes Factor will favour the unconstrained model $H_u$.

>>>>

\title{Evaluating Informative Hypotheses with Equality and Inequality Constraints:\\ A Tutorial Using the Bayes Factor via the Encompassing Prior Approach}
\shorttitle{The Bayes Factor via the Encompassing Prior Approach}

@@

\abstract{When conducting a study, researchers usually have expectations based on hypotheses or theoretical perspectives they want to evaluate. Equality and inequality constraints on the model parameters are used to formalize researchers' expectations or theoretical perspectives into the so-called informative hypotheses. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are unsuitable for testing complex informative hypotheses. An alternative approach is to use the Bayes factor. In particular, the Bayes factor based on the encompassing prior approach allows researchers to easily evaluate complex informative hypotheses in a wide range of statistical models (e.g., generalized linear). This paper provides a detailed introduction to the Bayes factor with encompassing prior. First, all steps and elements involved in the formalization of  informative hypotheses and the computation of the Bayes factor with encompassing prior are described. Next, we apply this method to a real case scenario, considering the attachment theory. Specifically, we analyzed the relative influence of maternal and paternal attachment on children's social-emotional development by comparing the various theoretical perspectives debated in the literature.

@@

The evaluation of hypotheses is one of the main goals of empirical research. In fact, when conducting a study, researchers usually want to evaluate the plausibility of specific hypotheses according to the collected data. These hypotheses can be formalized in accordance with researchers' expectations or theoretical perspectives into the so-called informative hypotheses. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are not suitable for evaluating the plausibility of complex informative hypotheses. Alternatively, the Bayes factor via the encompassing prior approach allows researchers to easily evaluate the plausibility of complex informative hypotheses in a wide range of statistical models. This paper provides a detailed introduction to the Bayes factor with encompassing prior. First, we describe all steps and elements involved in the formalization of  informative hypotheses and the computation of the Bayes factor. Next, we apply this method to a real case scenario, considering the attachment theory. Specifically, we analyzed the relative influence of maternal and paternal attachment on children's social-emotional development by comparing the various theoretical perspectives debated in the literature. This approach is of particular interest in applied research, as it allows researchers to formalize hypotheses with great flexibility and to directly evaluate which hypothesis is the most supported by the data.
}

@@

The dominant statistical approach to evaluate research hypotheses in psychology is the Null Hypothesis Significance Testing (NHST). In the literature, however, the NHST's utility and validity are largely debated \parencite{wassersteinMovingWorld052019}. This approach presents several limitations. First, the NHST places a narrow focus on statistical testing rather than on the formalization of hypotheses, leading researchers to evaluate data against the catch-all null hypothesis that nothing is going on rather than testing their specific expectation (and the alternative hypothesis is rarely formalized). Second, the \textit{p}-value does not quantify the evidence in favor of one hypothesis; therefore, it is not possible to \textit{“accept”} a hypothesis but only to \textit{“reject”} it. This situation has some drawbacks because nonsignificant results leave researchers in a state of indecision. Third, the NHST does not allow researchers to test multiple hypotheses at the same time. With the NHST, the null hypothesis is tested only against a single alternative. Fourth, the NHST is unsuitable for testing broad classes of hypotheses with equality and inequality constraints \parencite{mulderSimpleBayesianTesting2019}. Researchers can evaluate expectations using one-side tests; however, when more groups or more variables are involved, it is not possible to evaluate complex parameter constraints that reflect researchers' expectations \parencite{vandeschootIntroductionBayesianModel2011}.

Model comparison is a different approach that allows researchers to compare multiple hypotheses and identify the most supported by the data \parencite{mcelreathStatisticalRethinkingBayesian2020}. Researchers first formalize hypotheses as statistical models according to their expectations or theoretical perspectives. Therefore, it is possible to determine which is the most supported model among those considered according to the data. To do so, a popular approach is to use information criteria, such as the AIC or BIC criteria, that estimate a model's ability to predict new data \parencite{wagenmakersAICModelSelection2004,akaike1973a, schwarzEstimatingDimensionModel1978}. Model comparison has several advantages; in particular, it allows researchers to directly compute each model's relative plausibility given the data and the set of models considered. However, model comparison using information criteria is not appropriate to evaluate hypotheses that include complex parameter constraints reflecting researchers' expectations. In fact, information criteria evaluate models' complexity according to the number of parameters, but they do not take into account possible order constraints on the parameters.

An alternative criterium to evaluate research hypotheses in a model comparison is the Bayes factor. During the last 25 years, interest in the Bayes factor has increased, and its use has been proposed as the solution to the NHST's critical issues \parencite{heckReviewApplicationsBayes2020, mulderEditorsIntroductionSpecial2016}. Although it has its own limitations and it does not solve the fundamental issues of the misuse of statistical techniques \parencite{schadWorkflowTechniquesRobust2021, gelmanBayesianDataAnalysis2013}, the Bayes factor offers some clear advantages. In particular, it allows us to compare hypotheses obtaining a relative index of evidence, like the information criteria, but it also allows us to easily compare complex research hypotheses. In fact, researchers can formalize so-called \textit{“informative hypotheses”}, hypotheses containing information about the ordering of the model parameters,  according to their expectations or theoretical perspectives, using equality and inequality constraints \parencite{vandeschootIntroductionBayesianModel2011}. Researchers can then compare these hypotheses against each other, using the Bayes factor.

The evaluation of informative hypotheses is of particular interest to researchers, as it allows them to assess specific complex expectations and theoretical perspectives directly. In the literature, a particular approach allowing researchers to easily compute the Bayes factor with complex informative hypotheses has received increasing attention: \textit{“the Bayes factor with encompassing prior”}. \textcite{vandeschootIntroductionBayesianModel2011} presented a general introduction to informative hypothesis testing using the Bayes factor with the encompassing prior approach whereas \textcite{hoijtinkInformativeHypothesesTheory2012} offered a more detailed description of this method's development. In other studies, however, researchers considered the application of this approach with specific statistical models, such as mixed-effect models \parencite{katoBayesianApproachInequality2006}, evaluation of correlation coefficients \parencite{mulderBayesFactorsTesting2016}, and multinomial models \parencite{heckMultinomialModelsLinear2019}. In addition, \textcite{guBayesianEvaluationInequality2014} proposed a general approximate procedure to evaluate inequality constraints in a wide range of statistical models.

However, in all the studies mentioned above, researchers considered only informative hypotheses with inequality constraints. To evaluate equality and inequality constraints in the same hypothesis, researchers had to approximate equality constraints to “\textit{about equality constraints}” (i.e., a equality constraints of type $\theta_i = \theta_j$ is approximated to $|\theta_i - \theta_j| < \xi$ for a value of $\xi$ small enough).

Only recently, \textcite{guApproximatedAdjustedFractional2018} introduced an approximate procedure to evaluate equality and inequality constraints, using the Bayes factor with the encompassing prior approach in a wide range of statistical models (i.e., generalized linear mixed models and structural equation models). Subsequently, \textcite{mulderBayesFactorTesting2019} extended this approach to generalized multivariate probit models, and, finally, \textcite{mulderSimpleBayesianTesting2019} proposed an accurate procedure (i.e., not based on an approximation) that allows for testing of informative hypotheses with equality and inequality constraints in linear regression models.

The development and implementation of this approach are of particular interest because although it has limits, it allows researchers to directly evaluate their expectations and hypotheses. The available literature, however, is rather technical. These articles' complexity makes it difficult for researchers not familiar with this approach to clearly understand all the steps involved. On the other hand, articles offering a general introduction to the Bayes factor do not provide enough details to allow readers to autonomously apply this method to their problems; instead, they usually rely on ad hoc solutions implemented in some statistical software. For example, \textcite{hoijtinkTutorialTestingHypotheses2019a} offered an excellent introduction to the Bayes factor for the evaluation of informative hypotheses using the R package \texttt{bain}. However, it is limited to the context of analysis-of-variance models, and it is not possible to extend its use to other cases.

The aim of this paper, therefore, is to offer a clear and detailed description of the Bayes factor with the encompassing prior approach to allow other researchers to apply this approach in their studies. In particular, we refer to the approximated method \textcite{guApproximatedAdjustedFractional2018} proposed, as it applies to a wider range of conditions than the more accurate approach \textcite{mulderSimpleBayesianTesting2019} introduced. In particular, researchers can use the proposed method in the case of generalized linear mixed models and structural equation models, two very general families of statistical models. Moreover, by providing an exhaustive description of all of this approach's steps, researchers do not have to rely on already available solutions; instead, they can implement this approach according to their specific needs. The paper is organized as follows. First, we introduce the method, providing a detailed description of all steps and elements involved in the formalization of informative hypotheses and Bayes factor computation. We then present an application of the method to real data to discuss the typical issues encountered in real complex scenarios.

@@

The evaluation of informative hypotheses with equality and/or inequality constraints involves several steps and elements. In this section, first, we describe how to formulate informative hypotheses. We then introduce the Bayes factor, considering the encompassing prior approach based on the approximated method \textcite{guApproximatedAdjustedFractional2018} proposed.

@@

\subsection{Formulation of Informative Hypotheses}\label{sec:informative-hypotheses}

Researchers can define informative hypotheses according to their expectation, evidence from the literature, or theoretical perspectives, and they form them with equality and/or inequality constraints on certain model parameters. These constraints are obtained as a linear combination of certain parameters and eventual constant values. For example, it is possible to state that two parameters are equal ($\theta_i = \theta_j$) or that one parameter is greater than another ($\theta_i > \theta_j$) or to express other complex conditions, such as the difference between two parameter is less than a given value ($\theta_i - \theta_j < .5$; $2\times\theta_i - \theta_j < .1 \times \sigma$).

Therefore, an informative hypothesis, $H_i$, with equality and inequality constraints can be expressed in the form
\begin{equation}
H_i:\ \ R_E\bm{\theta} = r_E \ ~ \ \& \ ~ \ R_I\bm{\theta} > r_I,
\end{equation}
where $R_E$ is a matrix expressing the equality constraints and $r_E$ is a vector containing the equality constraints' constant values. $R_I$ is a matrix expressing the inequality constraints, and $r_E$ is a vector containing the inequality constraints' constant values. Finally, $\bm{\theta}$ is a vector with the model parameters involved in the constraints.

As an example, consider a study evaluating the efficacy of a new psychological treatment that is supposed to improve a given cognitive ability. In the study, a control group receiving no treatment and another group receiving the traditional treatment were included as a comparison. Moreover, imagine that the new psychological treatment was administered in three different modalities to different groups: individually, in-group, and online. Researchers expect no differences between the three modalities, but they hypothesize that the new treatment will perform better than the traditional one, and this result, in turn, will be better than the no-treatment control condition. We can express this hypothesis as
\begin{equation}
H_i:\ \ \theta_{control}< \theta_{traditional} < \theta_{individual} = \theta_{group} = \theta_{online},
\end{equation}
were $\theta_{control}$ is the parameter estimating the average score of the no-treatment control group, $\theta_{traditional}$ is the parameter estimating the average score of the group receiving the traditional treatment, and $\theta_{individual}$, $\theta_{group}$, and $\theta_{online}$ are the parameters estimating the average scores of the groups receiving the new treatment individually, in-group, and online, respectively. The corresponding formulation of the hypothesis using the matrix notation previously introduced is
\begin{equation}
\begin{aligned}
H_i: &\\
& R_E\bm{\theta} =
\begin{bmatrix}
0 & 0 & 1 & -1 & 0\\
0 & 0 & 0 & 1 & -1
\end{bmatrix}        \begin{bmatrix} \theta_{control} \\ \theta_{traditional} \\ \theta_{individual} \\ \theta_{group} \\ \theta_{online}\end{bmatrix} = \begin{bmatrix} 0 \\ 0\end{bmatrix} = r_E,\\~\\
& R_I\bm{\theta} =
\begin{bmatrix}
-1 & 1 & 0 & 0 & 0\\
0 & -1 & 1 & 0 & 0
\end{bmatrix}        \begin{bmatrix} \theta_{control} \\ \theta_{traditional} \\ \theta_{individual} \\ \theta_{group} \\ \theta_{online}\end{bmatrix} > \begin{bmatrix} 0 \\ 0\end{bmatrix} = r_I.
\end{aligned}
\end{equation}
Note how each row of $R_E$ and $R_I$ matrices expresses an equality or an inequality constraint, respectively. For example, in the first row of $R_E$, we have $\theta_{individual} - \theta_{group} = 0$ (i.e., $\theta_{individual} = \theta_{group}$) and in the first row of $R_I$ we have $\theta_{traditional} - \theta_{control} > 0$ (i.e., $\theta_{traditional} > \theta_{control}$).

Now that we have explained how to define an informative hypothesis with equality and inequality constraints introducing an appropriate notation, let's see how to evaluate an informative hypothesis using the Bayes factor with the encompassing prior approach.

@@

The Bayes factor of hypothesis $H_1$ against a competing hypothesis, $H_2$, is defined as the ratio between the marginal likelihoods of the two hypotheses:
\begin{equation}
BF_{12} = \frac{Pr(Y|H_1)}{Pr(Y|H_2)} = \frac{\int l(Y|\theta_1, H_1) \pi(\theta_1| H_1)\,d\theta_1}{\int l(Y|\theta_2, H_2) \pi(\theta_2| H_2)\,d\theta_2},
\end{equation}
where $Y$ indicates the data, $\theta_i$ is the vector of parameters under the hypothesis $H_i$, $l(Y|\theta_i, , H_i)$ is the likelihood function under the hypothesis $H_i$, and $\pi(\theta_i| H_i)$ is the prior of the parameters under the hypothesis $H_i$. The marginal likelihood $Pr(Y|H_i)$ can be interpreted as a measure of the data's plausibility under $H_i$.

Therefore, the Bayes factor, $BF_{12}$, quantifies the relative support of the data for the two competing hypotheses (and not the ratio between the probability of the two hypotheses). For values close to 1, the Bayes factor indicates that $H_1$ and $H_2$ have similar support from the data. Larger values indicate evidence in favor of $H_1$ whereas values close to 0 indicate evidence in favor of $H_2$. The ratio between these two hypotheses' posterior probabilities can be computed as
\begin{equation}
\frac{Pr(H_1|Y)}{Pr(H_2|Y)} = \overbrace{\frac{Pr(Y|H_1)}{Pr(Y|H_2)}}^{BF_{12}} \times \frac{Pr(H_1)}{Pr(H_2)},
\end{equation}
where $Pr(H_i)$ is the prior probability of $H_i$ and $Pr(H_i|Y)$ is the posterior probability of $H_i$. For a detailed description of the Bayes factor, considering also its interpretation and application in various contexts see \textcite{wagenmakersBayesianHypothesisTesting2010, mulderEditorsIntroductionSpecial2016, heckReviewApplicationsBayes2020}.

Note that to compute the marginal likelihood $Pr(Y|H_i)$ of a hypothesis, $H_i$, it is necessary to integrate the product between the likelihood and the prior. These integrals, however, are usually difficult to compute (in particular when hypotheses include order constraints). In the case of hypotheses with equality and inequality constraints, however, it is possible to simplify the computation of the Bayes factor using the encompassing prior approach.

@@

The basic idea of the encompassing prior approach is to consider an informative hypothesis as a subset of an unconstrained model's parameter space. Therefore, to evaluate a hypothesis's plausibility, we can consider the proportion of parameter space of the unconstrained model that satisfies the constraints. More specifically, given an informative hypothesis, $H_i$, and an unconstrained model, $H_u$, (or \textit{encompassing model}) that does not contain any constraints on the parameters, if the prior under $H_i$ is defined as a truncation of the proper prior under $H_u$ (\textit{encompassing prior}) according to the constraints, then the Bayes factor between $H_i$ and $H_u$ can be written as
\begin{equation}\label{eq:BF-encompassing}
\begin{aligned}
BF_{iu} &= \frac{Pr(\text{Inequality Const|Equality Const, Data, } H_u)}{\pi(\text{Inequality Const|Equality Const, } H_u)} \times \frac{Pr(\text{Equality Const|Data, } H_u)}{\pi(\text{Equality Const}|H_u)} \\ ~\\
&= \frac{Pr(R_I\theta > r_I | R_E\theta =r_E, Y, H_u)}{\pi(R_I\theta > r_I | R_E\theta =r_E, H_u)} \times \frac{Pr(R_E\theta =r_E| Y, H_u)}{\pi(R_E\theta =r_E| H_u)}.
\end{aligned}
\end{equation}

The first term is the ratio between the conditional posterior probability and the conditional prior probability that the inequality constraints hold under the unconstrained model, $H_u$, given the equality constraints. The second term is the ratio between marginal posterior density and the marginal prior density of the equality constraints under $H_u$ \parencite[the well-known Savage–Dickey density ratio; ][]{dickeyWeightedLikelihoodRatio1971, wetzelsEncompassingPriorGeneralization2010}. In particular, the four elements can be interpreted as:
\begin{itemize}
  \item{The \textbf{conditional posterior probability}, $Pr(R_I\theta > r_I | R_E\theta =r_E, Y, H_u)$, is a measure of the fit of the inequality constraints of $H_i$ under $H_u$.}
  \item{The \textbf{conditional prior probability}, $\pi(R_I\theta > r_I | R_E\theta =r_E, H_u)$, is a measure of the complexity of the inequality constraints of $H_i$ under $H_u$.}
  \item{The \textbf{marginal posterior density}, $Pr(R_E\theta =r_E| Y, H_u)$, is a measure of the fit of the equality constraints of $H_i$ under $H_u$.}
  \item{The \textbf{marginal prior density}, $\pi(R_E\theta =r_E| H_u)$, is a measure of the complexity of the equality constraints of $H_i$ under $H_u$.}
\end{itemize}
Summarizing at the denominators, we have measures of the \textbf{complexity} of the inequality and equality constraints of the informative hypothesis $H_i$; at the numerators, instead, we have measures of the \textbf{fit} of the data to the inequality and equality constraints of the informative hypothesis $H_i$. Therefore, if the hypothesis $H_i$, although applying constraints to the parametric space (less complexity), can still provide an accurate description of the data, the Bayes factor will favor $H_i$. On the contrary, if the support of the data is poor, the Bayes factor will favor the unconstrained model, $H_u$.
